---
title: "Analysis Scripts"
author: "Yi Zhang"
date: "4/24/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load-pkg, include = FALSE, message = FALSE}
library(ggplot2)  # for plots
library(magrittr)  # for `%>%` operator
library(haven)  # for reading sav files
library(modelsummary)  # for summarizing data
library(rstan)
rstan_options(auto_write = TRUE)  # save compiled STAN object
options(mc.cores = 2)  # use two cores
library(brms)
library(posterior)
library(bayesplot)
library(kableExtra)
theme_set(theme_classic() +
    theme(panel.grid.major.y = element_line(color = "grey92")))
```
## PSYC-573 Final Project Analysis Scripts

### Cleaning data and summarize variables of interest
```{r clean data, echo = FALSE}
# Import data
morality_data <- read_sav("/Users/yizhang/Desktop/USC/Coursework/Spring22/Bayesian/PSYC-573 Project/data/Study4_FinalData.sav")
morality_data$Group <- as.factor(morality_data$Group)
#exclude unqualified subjects
morality_data <- morality_data[morality_data$`filter_$` == 1, ]
#summarize data
datasummary((Other1Moral_GC + Other2Moral + SCEmo)  * 
                (N + Mean + SD + Min + Max) ~ 
                factor(Group, labels = c("Romantic Partner", "Friend", "Stranger")),
            data = morality_data,
            # The `caption` argument adds caption to a table.
            caption = "Summary Statistics") %>%
    # add table note
    add_footnote("SD = standard deviation", notation = "none")

#plot DVs by condition
par(mfrow = c(1,3))
#moral judgment (baseline)
ggplot(morality_data, aes(x = Other1Moral)) +
    geom_histogram(bins = 10) +
    facet_wrap(~ Group,  ncol = 1) +
    labs(x = "Moral judgment after learning transgression")
#moral judgment (after learning)
ggplot(morality_data, aes(x = Other2Moral)) +
    geom_histogram(bins = 10) +
    facet_wrap(~ Group,  ncol = 1) +
    labs(x = "Moral judgment after learning transgression")
#self-conscious emotion
ggplot(morality_data, aes(x = SCEmo)) +
    geom_histogram(bins = 10) +
    facet_wrap(~ Group,  ncol = 1) +
    labs(x = "Self-conscious emotions after learning transgression")
```
### Model equations and priors

#### Moral judgment
$$
  \begin{aligned}
    \text{Moral_Rating}_{ijk} \sim \beta_{0jk} + \beta_{1j}\text{Baseline_GC}\\
    \beta_{0jk} \sim N(\mu_j, \sigma_1)\\
    \beta_j \sim N(\beta_k, \sigma_2)\\
    \beta_k \sim N(\beta_0, \sigma_3)\\
    \beta_{1j} \sim N(\beta_1, \sigma_{21})\\\\
    \sigma_1 \sim t^+_4(0,5)\\
    \sigma_2 \sim t^+_4(0,5)\\
    \sigma_3 \sim t^+_4(.3,1)\\
    \beta_0 \sim t^+_4(3,2)\\
    \beta_1 \sim t^+_4(0,5)\\
    \sigma_{21} \sim t^+_4(0,5)\\
    \rho \sim \text{LKJ}(1)
  \end{aligned}
$$
#### Self-conscious emotions
$$
  \begin{aligned}
    \text{SCEmo}_{ijk} \sim \text{LogNormal}(\mu_{jk}, \sigma_1)\\
    \mu_{jk} \sim N(\mu_{k}, \sigma_{2})\\
    \mu_k \sim N(\mu_{0}, \sigma_{3})\\\\
    \sigma_{1} \sim t^+_4(0,5)\\
    \sigma_{2} \sim t^+_4(1,1)\\
    \sigma_{3} \sim t^+_4(.5, 1)\\
    \mu_{0} \sim t^+_4(2.5,2)
  \end{aligned}
$$

### Moral judgment by condition (controlling for baseline)
#### Running model usng `brms`
```{r running model, echo=TRUE}
#get rid of NAs
morality_data.noNA <- morality_data[!is.na(morality_data$Other2Moral) &
                                      !is.na(morality_data$Other1Moral_GC),]

m1_brm <- brm(
    Other2Moral ~ Other1Moral_GC + (1|Group) + (Other1Moral_GC|Group:Pair_ID),
    data = morality_data.noNA,
    # Normal distribution with identity link
    family = gaussian(link = "identity"),
    # Overwrite the default priors
    prior = c(
        # prior for intercept (default)
        prior(student_t(4, 3, 2), class = "Intercept"),
        # prior for beta1 (and beta2, ..., if > 1 predictors)
        # prior for fixed effect of covariate
        prior(student_t(4, 0, 5), class = "b", coef = "Other1Moral_GC"),
        # prior for sd (between groups)
        prior(student_t(4, .3, 1), class = "sd", coef = "Intercept", group = "Group"),
        # prior for sd (between-observations)
        prior(student_t(4, 0, 5), class = "sigma"),
        # correlation
        prior(lkj(1), class = "cor")
    ),
    control = list(adapt_delta = 0.9999, max_treedepth = 10),
    sample_prior = TRUE,  # also sample the prior distributions
    iter = 10000,  # default is 4 chains, 2000 iterations each
    seed = 1234
)
#examine output
summary(m1_brm)

#create draws array and summarize results
#compare groups
coef <- coef(m1_brm)

#mutate variables
draws <- as.array(m1_brm, variable = c("b_Intercept",
                      "b_Other1Moral_GC",
                      "sigma",
                      "r_Group[1,Intercept]",
                      "r_Group[2,Intercept]",
                      "r_Group[3,Intercept]")) %>%
    as_draws() %>%
    mutate_variables(friend_minus_romantic = `r_Group[2,Intercept]` - 
                       `r_Group[1,Intercept]`,
                     stranger_minus_romantic = `r_Group[3,Intercept]` - 
                       `r_Group[1,Intercept]`,
                     stranger_minus_friend = `r_Group[3,Intercept]` - 
                       `r_Group[2,Intercept]`)
draws_summary <- draws%>%
    summarize_draws()
#results show that group 3 is lower than both group 1 and group 2
#but there's no difference between group 1 and group 2
```

#### Checking model convergence
```{r check convergence, echo=TRUE}
#rank histograms
mcmc_rank_hist(draws, pars = 
                 c("b_Intercept",
                      "b_Other1Moral_GC",
                      "sigma",
                      "r_Group[1,Intercept]",
                      "r_Group[2,Intercept]",
                      "r_Group[3,Intercept]"))
#trace plot
mcmc_trace(m1_brm,
           pars = c("b_Intercept",
                      "b_Other1Moral_GC",
                      "sigma",
                      "r_Group[1,Intercept]",
                      "r_Group[2,Intercept]",
                      "r_Group[3,Intercept]"))
```

#### Plotting model results
```{r plot results, echo = TRUE}
#visualizing
mcmc_areas(draws, pars = c("r_Group[1,Intercept]",
                      "r_Group[2,Intercept]",
                      "r_Group[3,Intercept]"))
mcmc_areas(draws, pars = c("friend_minus_romantic", 
                            "stranger_minus_romantic",
                            "stranger_minus_friend"))

#model summary table
msummary(m1_brm, statistic = "conf.int", fmt = 2)
```

#### Posterior predictive check
```{r ppc, echo = TRUE}
pp_check(
    m1_brm,
    type = "intervals", 
    x = "Other1Moral_GC"
) +
    labs(x = "Baseline morality rating", y = "Morality rating after learning transgression")

#density plot overlay
pp_check(m1_brm, type = "dens_overlay_grouped", group = "Group", ndraws = 20)
```

### Self-conscious emotions by condition
```{r running second model in brms, echo = TRUE}
morality_data.noNA2 <- morality_data[!is.na(morality_data$SCEmo),]
m2_brm <- brm(
    SCEmo ~ (1|Group) + (1|Group:Pair_ID),
    data = morality_data.noNA2,
    # Normal distribution with identity link
    family = lognormal(link = "identity"),
    # Overwrite the default priors
    prior = c(
        # prior for intercept
        prior(student_t(4, 2.5, 2), class = "Intercept"),
        # prior for sd (between groups)
        prior(student_t(4, .5, 1), class = "sd", coef = "Intercept", group = "Group"),
        # prior for sd (between dyads)
        prior(student_t(4, 1, 1), class = "sd", coef = "Intercept", group = "Group:Pair_ID"),
        # prior for sigma (sd between observations)
        prior(student_t(4, 0, 5), class = "sigma")
    ),
    control = list(adapt_delta = .99995, max_treedepth = 12),
    sample_prior = TRUE,  # also sample the prior distributions
    iter = 10000,  # default is 4 chains, 2000 iterations each
    seed = 1234
)

#examine output
summary(m2_brm)

coef2 <- coef(m2_brm)

#summarize draws
draws2 <- as.array(m2_brm, variable = c("b_Intercept",
                      "sigma",
                      "r_Group[1,Intercept]",
                      "r_Group[2,Intercept]",
                      "r_Group[3,Intercept]")) %>%
    as_draws() %>%
    mutate_variables(friend_minus_romantic = `r_Group[2,Intercept]` - 
                       `r_Group[1,Intercept]`,
                     stranger_minus_romantic = `r_Group[3,Intercept]` - 
                       `r_Group[1,Intercept]`,
                     stranger_minus_friend = `r_Group[3,Intercept]` - 
                       `r_Group[2,Intercept]`)
draws_summary2 <- draws2%>%
    summarize_draws()
```

#### Checking model convergence (SCEmo)
```{r check convergence 2, echo=TRUE}
#rank histograms
mcmc_rank_hist(draws2, pars = 
                 c("b_Intercept",
                      "sigma",
                      "r_Group[1,Intercept]",
                      "r_Group[2,Intercept]",
                      "r_Group[3,Intercept]"))
#trace plot
mcmc_trace(m2_brm,
           pars = c("b_Intercept",
                      "sigma",
                      "r_Group[1,Intercept]",
                      "r_Group[2,Intercept]",
                      "r_Group[3,Intercept]"))
```

#### Plotting model results (SCEmo)
```{r plot results 2, echo = TRUE}
#visualizing
mcmc_areas(draws, pars = c("r_Group[1,Intercept]",
                      "r_Group[2,Intercept]",
                      "r_Group[3,Intercept]"))
mcmc_areas(draws2, pars = c("friend_minus_romantic", 
                            "stranger_minus_romantic",
                            "stranger_minus_friend"))
#model summary table
msummary(m2_brm, statistic = "conf.int", fmt = 2)
```

#### Posterior predictive check (SCEmo)
```{r ppc, echo = TRUE}
#density plot overlay
pp_check(m2_brm, type = "dens_overlay_grouped", group = "Group", ndraws = 20)
```