---
title: "HW 9"
shorttitle: "A Bayesian replication of Forbes & Stellar (2021)"
author:
  - name: "Yi Zhang"
    affiliation: "1"
affiliation:
  - id: "1"
    institution: "University of Southern California"
authornote:
abstract:
header-includes:
  - \usepackage{siunitx}
csl: "`r system.file('rmd', 'apa7.csl', package = 'papaja')`"
documentclass: "apa7"
output: papaja::apa6_pdf
bibliography: "references.bib"
---

<!-- This is a comment -->
<!-- To use the `papaja::apa6_pdf` format, follow the instruction in https://github.com/crsh/papaja to install the `papaja` package -->

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r load-pkg, include = FALSE, message = FALSE}
library(papaja)
library(ggplot2)  # for plots
library(magrittr)  # for `%>%` operator
library(here)
library(haven)  # for reading sav files
library(modelsummary)  # for summarizing data
library(rstan)
rstan_options(auto_write = TRUE)  # save compiled STAN object
options(mc.cores = 2)  # use two cores
library(brms)
library(posterior)
library(bayesplot)
library(kableExtra)
theme_set(theme_classic() +
    theme(panel.grid.major.y = element_line(color = "grey92")))
```

# Introduction

## Research Question

> How does relationship context affect moral judgment and moral emotion?

# Method

> Each participant was randomly assigned to learn about moral transgressions committed by either 1) their romantic partner, 2) a close friend, or 3) a stranger. Participants first rated their partner's morality and then learned about moral transgressions ostensibly committed by their partner. Finally, participants rated their partner's morality again and reported their own emotions.

## Variables

- `Pair_ID`: ID for each dyad (romantic partners, friends, or strangers)
- `Group`: Condition. 1: Romantic Partner, 2: Close friend, 3: Stranger
- `Other1Moral`: Morality rating at baseline. "After learning this information, please rate the other person on the following traits. (honest, trustworthy, fair, unethical)" 1 (Not at all) to 5 (Extremely)
- `Other2Moral`: Morality rating after learning about transgression. "After learning this information, please rate the other person on the following traits. (honest, trustworthy, fair, unethical)" 1 (Not at all) to 5 (Extremely)
- `SCEmo`: Self-conscious emotions after learning about transgression. "After learning this information, how much do you feel the following emotions your self? (Embarrassed, Shame, Guilt)" 1: None at all, 10: As much as I've ever felt.


```{r import data}
# Import data
morality_data <- read_sav("/Users/yizhang/Desktop/USC/Coursework/Spring22/Bayesian/PSYC-573 Project/data/Study4_FinalData.sav")
#exclude unqualified subjects
morality_data$Group <- as.factor(morality_data$Group)
morality_data <- morality_data[morality_data$`filter_$` == 1, ]
```

## Variable Summary

<!-- You can use \@ref(tab:xx) to reference a table with the chunk label "xx" -->

Table\ \@ref(tab:summ-var) shows the summary statistics of moral evaluations and self-conscious emotion ratings by conditions.

```{r summ-var}
datasummary((Other1Moral + Other2Moral + SCEmo)  * 
                (N + Mean + SD + Min + Max) ~ 
                factor(Group, labels = c("Romantic Partner", "Friend", "Stranger")),
            data = morality_data,
            # The `caption` argument adds caption to a table.
            caption = "Summary Statistics") %>%
    # add table note
    add_footnote("SD = standard deviation", notation = "none")
```

## Model

Let $Y$ = `Other2Moral`, `SCEmo`, $G$ = `Group`

For each of the DVs, I fit the following model:

$$
  \begin{aligned}
    Y_{j, G = romantic} & \sim N(\theta_{1j}, \sigma) \\
    Y_{j, G = friend} & \sim N(\theta_{2j}, \sigma)\\
    Y_{j, G = stranger} & \sim N(\theta_{3j}, \sigma)\\
    \theta_{1j}& \sim N(\mu_1, \sigma_1)\\
    \theta_{2j} & \sim N(\mu_2, \sigma_2)\\
    \theta_{3j} & \sim N(\mu_3, \sigma_3)
  \end{aligned}
$$

Below are the priors for the `OtherMoralDif` variable, based on data from Study 1.
$$
  \begin{aligned}
    \mu_1 & \sim N(2.61,2)\\
    \mu_2 & \sim N(2.52,2)\\
    \mu_2 & \sim N(2.11,2)\\
    \sigma & \sim N^+(0, 2)\\
    \sigma_1 & \sim N^+(1.6, 2) \\
    \sigma_2 & \sim N^+(1.35, 2)\\
    \sigma_2 & \sim N^+(1.35, 2)
  \end{aligned}
$$

Below are the priors for the `SCEmo` variable.
$$
  \begin{aligned}
    \mu_1 & \sim N(2.89,2)\\
    \mu_2 & \sim N(2.42,2)\\
    \mu_2 & \sim N(1.85,2)\\
    \sigma & \sim N^+(0, 2)\\
    \sigma_1 & \sim N^+(1.17, 2) \\
    \sigma_2 & \sim N^+(1.07, 2)\\
    \sigma_2 & \sim N^+(1.06, 2)
  \end{aligned}
$$

## Running Stan

We ran a separate model for each DV. In each model, we used 4 chains, each with 4,000 iterations (first 2,000 as warm-ups). For this preliminary analysis, I did not control for baseline moral judgment (`Other2Moral`). I also treated participants in the same dyad individuals, without taking into account the within-dyadic correlations.

```{r m1, results = "hide"}
#get rid of NA
morality_data.noNA <- morality_data[!is.na(morality_data$Other2Moral), ]

# 1. form the data list for Stan
stan_dat <- with(morality_data.noNA,
    list(N1 = sum(Group == 1),
         N2 = sum(Group == 2),
         N3 = sum(Group == 3),
         y1 = Other2Moral[which(Group == 1)],
         y2 = Other2Moral[which(Group == 2)],
         y3 = Other2Moral[which(Group == 3)]
         )
)
# 2. Run Stan
m1 <- stan(
    file = "/Users/yizhang/Desktop/USC/Coursework/Spring22/Bayesian/PSYC-573 Project/hw9.stan",
    data = stan_dat,
    seed = 1234,  # for reproducibility
    iter = 4000
)
```

```{r m2, results = "hide"}
#get rid of NA
morality_data.noNA2 <- morality_data[!is.na(morality_data$SCEmo), ]

# 1. form the data list for Stan
stan_dat2 <- with(morality_data.noNA2,
    list(N1 = sum(Group == 1),
         N2 = sum(Group == 2),
         N3 = sum(Group == 3),
         y1 = SCEmo[which(Group == 1)],
         y2 = SCEmo[which(Group == 2)],
         y3 = SCEmo[which(Group == 3)]
         )
)
# 2. Run Stan
m2 <- stan(
    file = "/Users/yizhang/Desktop/USC/Coursework/Spring22/Bayesian/PSYC-573 Project/hw9_2.stan",
    data = stan_dat2,
    seed = 1234,  # for reproducibility
    iter = 4000
)
```

```{r m1 brms, eval = FALSE, include = FALSE}
#Fitting model using brms (NEED TO REVISE CODE)
morality_data.noNA <- morality_data[!is.na(morality_data$Other2Moral) &
                                      !is.na(morality_data$Other1Moral_GC),]

#create stan code
stan_code_test <- make_stancode(Other2Moral ~ Group + Other1Moral_GC + (1|Pair_ID),
                                data = morality_data.noNA)
#fitting model for each condition separately
stan_code_test <- make_stancode(Other2Moral ~ Other1Moral_GC + (1|Pair_ID),
                                data = morality_data.noNA%>%
                                  filter(Group==1))
#fit brm model
m1_brm <- brm(
    # Y (vote) = beta0 + beta1 (growth)
    Other2Moral ~ Other1Moral_GC + (1|Group),
    data = morality_data.noNA,
    # Normal distribution with identity link
    family = gaussian(link = "identity"),
    # Overwrite the default priors
    prior = c(
        # prior for beta0
        prior(normal(3, 2), class = "Intercept"),
        # prior for beta1 (and beta2, ..., if > 1 predictors)
        prior(normal(0, 1), class = "b"),
        # prior for sigma
        prior(student_t(4, 0, 5), class = "sigma")
    ),
    control = list(adapt_delta = 0.99),
    sample_prior = TRUE,  # also sample the prior distributions
    iter = 10000,  # default is 4 chains, 2000 iterations each
    seed = 1234
)

summary(m1_brm)
test_draws <- as_draws(m1_brm)
prior_summary(m1_brm)
mcmc_areas(m1_brm, pars = c("b_Group2","b_Group3","b_Other1Moral_GC"))
```

# Results

<!-- Use [@xxx] for parenthetical citation. xxx is the label for the bibTeX citation key in "references.bib." See https://bookdown.org/yihui/rmarkdown-cookbook/bibliography.html for more information. -->
<!-- You can use \@ref(fig:xx) to reference a figure with the chunk label "xx" -->

```{r rank-hist-m1, fig.cap = "Rank histograms showing convergence of the MCMC chains for moral judgment."}
mcmc_rank_hist(m1, pars = c("mu1", "mu2", "mu3", "sigma1", "sigma2", "sigma3"))
```

```{r rank-hist-m2, fig.cap = "Rank histograms showing convergence of the MCMC chains for self-conscious emotions."}
mcmc_rank_hist(m2, pars = c("mu1", "mu2", "mu3", "sigma1", "sigma2", "sigma3"))
```

As shown in the rank histograms [@vehtari2021] in Figure\ \@ref(fig:rank-hist-m1) and Figure\ \@ref(fig:rank-hist-m2), the chains mixed well for both DVs.

Table\ \@ref(tab:summ-m1) shows the posterior distributions of $\mu_1$, $\mu_2$, $\mu_3$, $\sigma_1$, $\sigma_2$, $\sigma_3$ for `Other2Moral` and the difference between each combinations of the three groups.

Table\ \@ref(tab:summ-m2) shows the posterior distributions of $\mu_1$, $\mu_2$, $\mu_3$, $\sigma_1$, $\sigma_2$, $\sigma_3$ for `SCEmo` and the difference between each combinations of the three groups.

```{r summ-m2}
summ_m1 <- as_draws_df(m1) %>%
    subset_draws(variable = c("mu1", "mu2", "mu3", "sigma1", "sigma2", "sigma3")) %>%
    mutate_variables(`mu2 - mu1` = mu2 - mu1, #difference between group 2 and group 1
                     `mu3 - mu1` = mu3 - mu1, #dif between group 3 and group 1
                     `mu3 - mu2` = mu3 - mu2) %>% #dif between group 3 and group 2
    summarise_draws()
#summ_m1 <- as_draws_df(m1_brm) %>%
#    subset_draws(variable = c("b_Intercept", "b_Group2", "b_Group3", 
#                              "b_Other1Moral_GC", "sigma")) %>%
#    summarise_draws()
# Use `papaja::apa_table()` for APA formatted tables
apa_table(summ_m1, digits = 2,
          # table caption
          caption = "Posterior summary of model parameters for moral judgment.",
          # table note
          note = "sd = standard deviation. ess = effective sample size.")
```

```{r summ-m1}
summ_m2 <- as_draws_df(m2) %>%
    subset_draws(variable = c("mu1", "mu2", "mu3", "sigma1", "sigma2", "sigma3")) %>%
    mutate_variables(`mu2 - mu1` = mu2 - mu1, #difference between group 2 and group 1
                     `mu3 - mu1` = mu3 - mu1, #dif between group 3 and group 1
                     `mu3 - mu2` = mu3 - mu2) %>% #dif between group 3 and group 2
    summarise_draws()

apa_table(summ_m2, digits = 2,
          # table caption
          caption = "Posterior summary of model parameters for self-conscious emotions.",
          # table note
          note = "sd = standard deviation. ess = effective sample size.")
```

The analysis showed that on average, participants in the romantic partner and close friend groups made less negative evaluations for their partners compared with participants in the stranger group. The posterior mean of the difference in moral judgment between the romantic partner and stranger groups was `r round(summ_m1$mean[8], 2)` and a 90% CI of [`r round(summ_m1$q5[8], 2)`, `r round(summ_m1$q95[8], 2)`]. The posterior mean of the difference in moral judgment between the friend and stranger groups was `r round(summ_m1$mean[9], 2)` and a 90% CI of [`r round(summ_m1$q5[9], 2)`, `r round(summ_m1$q95[9], 2)`].

In the meantime, participants in the friend and stranger groups experienced less self-conscious emotions after learning that their partners had committed moral transgressions. The posterior mean of the difference between the romantic partner and friend groups was `r round(summ_m2$mean[7], 2)`, with a 90% CI of [`r round(summ_m2$q5[7], 2)`, `r round(summ_m2$q95[7], 2)`]. The posterior mean of difference between the romantic partner and stranger groups was `r round(summ_m2$mean[8], 2)`, with a 90% CI of [`r round(summ_m2$q5[8], 2)`, `r round(summ_m2$q95[8], 2)`].

\newpage

# References

::: {#refs custom-style="Bibliography"}
:::
